---
title: "STATS 271 Final Project"
author: 'Louise Huang (SUNet: qyhuang)'
date: "5/29/2020"
output: pdf_document
---

```{r}
library(mlmRev)
library(tidyverse)
library(rstan)
library(rstanarm)
library(gridExtra)
library(loo)
library(brms)
library(ggplot2)
theme_set(theme_minimal())
data(Contraception)
cont <- Contraception
#save.image("~/Desktop/271_project.RData")
load("~/Desktop/271_project.RData")
```

# 1. EDA and summary statistics

```{r}
use_cont <- filter(cont, use==1)
prop_one <- function(y) mean(y == 1)
group_cont_use <- cont %>% group_by(district) %>% summarise(use=mean(use))
group_region_use <- cont %>% group_by(urban) %>% summarise(use=mean(use))
table(cont$livch)
group_livch_use <- cont %>% group_by(livch) %>% summarise(use=mean(use))
mean(cont$use)
cont2 <- cont
cont2$age <- cont2$age + 30
test <- filter(cont2, age > 50 & age <= 55)
mean(test$use)

district_mean_hist <- group_cont_use %>% ggplot(aes(x=use)) + geom_histogram(binwidth=0.05, color="black", fill="white") + theme_bw() + ggtitle("Distribution of percentage of contraception use by district") + xlab("Percentage of contraception use") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

cont_2 <- use_cont %>% ggplot(aes(x = factor(urban), fill = factor(livch))) + geom_bar(position="dodge") + theme_bw() + ggtitle("Use of contraception by urban/rural residence and number of living children") + xlab("Urban Residence") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) + scale_fill_discrete(name="Number of living children")

grid.arrange(district_mean_hist, cont_2, ncol=2)
```



## Model 1. Varying intercept model with no predictor

```{r}
varying_inter_nopred_mod <- stan_glmer(formula = use ~ 1 + (1|district), 
                         data = cont,
                         seed = 271,
                         cores = getOption("mc.cores", 4L),
                         family = binomial(link = "logit"))
summary(varying_inter_nopred_mod,  digits = 3, 
        pars=c("(Intercept)", "b[(Intercept) district:1]", "b[(Intercept) district:2]", "Sigma[district:(Intercept),(Intercept)]"), 
        probs = c(0.025, 0.5, 0.975))
summary(varying_inter_nopred_mod)
prior_summary(varying_inter_nopred_mod)
```

```{r}
plot(varying_inter_nopred_mod, pars=c("(Intercept)", "b[(Intercept) district:1]", "b[(Intercept) district:2]", "Sigma[district:(Intercept),(Intercept)]"), 
     prob = 0.5, prob_outer = 0.95) + ggplot2::ggtitle("Posterior medians \n with 50% and 95% intervals in varying intercept model")
plot(varying_inter_nopred_mod, "trace", pars = "(Intercept)") + ggplot2::ggtitle("Traceplot of intercept in varying intercept model")
```

```{r}
yrep_mod1 <- posterior_predict(varying_inter_nopred_mod, draws = 2000)
ytrue <- cont$use
cont$use <- ifelse(cont$use=="Y", 1, 0) # convert Y/N to 1/0
mean(ytrue)
hist(apply(yrep_mod1, 1, mean))
```

```{r}
prop_zero_test1 <- pp_check(varying_inter_nopred_mod, plotfun = "stat", stat = "prop_one", binwidth=0.005) 
prop_zero_test1 + ggtitle("Posterior checking of varying intercept model with no predictors")
```

## Model 2. Varying intercept model with predictors
all the predictors have the same slope, but allows intercept to vary among districts.
```{r}
varying_intercept_mod <- stan_glmer(formula = use ~ 1 + urban + age + livch + (1|district), 
                         data = cont,
                         seed = 271,
                         cores = getOption("mc.cores", 4L),
                         family = binomial(link = "logit"))
prior_summary(varying_intercept_mod)
```

```{r}
summary(varying_intercept_mod)
summary(varying_intercept_mod,  digits = 3, 
        pars=c("(Intercept)", "b[(Intercept) district:1]", "b[(Intercept) district:2]", "urbanY", "age", "livch1", "livch2", "livch3+",
               "Sigma[district:(Intercept),(Intercept)]"), 
        probs = c(0.025, 0.5, 0.975))
```

Plotting
```{r}
plot(varying_intercept_mod, pars=c("(Intercept)", "b[(Intercept) district:1]", "b[(Intercept) district:2]", "urbanY", "age", "livch1", "livch2", "livch3+","Sigma[district:(Intercept),(Intercept)]")) + ggplot2::ggtitle("Posterior medians \n with 50% and 95% intervals")
plot(varying_intercept_mod, "trace", pars = "(Intercept)") + ggplot2::ggtitle("Traceplot of intercept")
```

```{r}
yrep_mod2 <- posterior_predict(varying_intercept_mod, draws = 2000)
mean(ytrue)
hist(apply(yrep_mod2, 1, mean))
```


```{r}
prop_zero_test2 <- pp_check(varying_intercept_mod, plotfun = "stat", stat = "prop_one", binwidth=0.005) 
prop_zero_test2 + ggtitle("Posterior checking of varying intercept model with predictors")
```

## Model 3. Varying intercept varying slope model
age and livch have the same slope, but allows the slope of urban and intercept to vary among the districts.
```{r}
varying_inter_slope_mod <- stan_glmer(formula = use ~ 1 + age + livch + (urban|district), 
                         data = cont,
                         seed = 271,
                         cores = getOption("mc.cores", 4L),
                         family = binomial(link = "logit"))
prior_summary(varying_inter_slope_mod)
plot(varying_inter_slope_mod)
```

```{r}
summary(varying_inter_slope_mod,  digits = 3, 
        pars=c("(Intercept)", "b[(Intercept) district:1]", "b[urbanY district:1]", "age", "livch1", "livch2", "livch3+","Sigma[district:(Intercept),(Intercept)]"), 
        probs = c(0.025, 0.5, 0.975))

summary(varying_inter_slope_mod)
```

```{r}
plot(varying_inter_slope_mod, pars=c("(Intercept)", "b[(Intercept) district:1]", "b[urbanY district:1]", "age", "livch1", "livch2", "livch3+","Sigma[district:(Intercept),(Intercept)]")) + ggplot2::ggtitle("Posterior medians \n with 50% and 95% intervals")
plot(varying_inter_slope_mod, "trace", pars = "(Intercept)") + ggplot2::ggtitle("Traceplot of intercept")
```


```{r}
plot(varying_inter_slope_mod, "trace", pars = "age")
```

Posterior checking
```{r}
yrep_mod3 <- posterior_predict(varying_inter_slope_mod, draws = 2000)
# mean(ytrue)
# hist(apply(yrep_mod3, 1, mean))

prop_one <- function(y) mean(y == 1)
prop_zero_test3 <- pp_check(varying_inter_slope_mod, plotfun = "stat", stat = "prop_one", binwidth=0.005) 
prop_zero_test3 + ggtitle("Posterior checking of varying intercept varying slope model")
```

```{r}
varying_inter_nopred_mod$waic <- rstanarm::waic(varying_inter_nopred_mod)
varying_intercept_mod$waic <- rstanarm::waic(varying_intercept_mod)
varying_inter_slope_mod$waic <- rstanarm::waic(varying_inter_slope_mod)

varying_inter_nopred_mod$loo <- rstanarm::loo(varying_inter_nopred_mod, cores = getOption("mc.cores", 4))
varying_intercept_mod$loo <- rstanarm::loo(varying_intercept_mod, cores = getOption("mc.cores", 4))
varying_inter_slope_mod$loo <- rstanarm::loo(varying_inter_slope_mod, cores = getOption("mc.cores", 4))

loo_compare(varying_inter_nopred_mod, varying_intercept_mod, varying_inter_slope_mod, criterion =  "loo")
loo_compare(varying_inter_nopred_mod, varying_intercept_mod, varying_inter_slope_mod, criterion =  "waic")

model_list <- stanreg_list(varying_inter_nopred_mod, varying_intercept_mod, varying_inter_slope_mod)
loo_model_weights(model_list)
```

Bayesian model averaging
```{r}
fit1 <- brm(formula = use ~ 1 + urban + age + livch + (1|district), 
                         data = cont,
                         seed = 271,
                         cores = getOption("mc.cores", 4L),
                         family = bernoulli(link = "logit"))
summary(fit1)

fit2 <- brm(formula = use ~ 1 + age + livch + (urban|district), 
                         data = cont,
                         seed = 271,
                         cores = getOption("mc.cores", 4L),
                         family = bernoulli(link = "logit"))
summary(fit2)
avg_pred <- pp_average(fit1, fit2)
```

